---
title: "Future50 Case Study"
author: "Josh Yednock"
date: "3/1/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_chunk$set(dev = 'pdf')

# Referencing packages I will be using -------------

if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(dplyr)) install.packages("dplyr", repos = "http://cran.us.r-project.org")
if(!require(readxl)) install.packages("readxl", repos = "http://cran.us.r-project.org")
if(!require(usmap)) install.packages("usmap", repos = "http://cran.us.r-project.org")
if(!require(ggplot2)) install.packages("ggplot2", repos = "http://cran.us.r-project.org")
if(!require(recipes)) install.packages("recipies", repos = "http://cran.us.r-project.org")
if(!require(forecast)) install.packages("forecast", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(scales)) install.packages("scales", repos = "http://cran.us.r-project.org")
if(!require(knitr)) install.packages("knitr", repos = "http://cran.us.r-project.org")
if(!require(vip)) install.packages("vip", repos = "http://cran.us.r-project.org")
if(!require(readxl)) install.packages("readxl", repos = "http://cran.us.r-project.org")
if(!require(GGally)) install.packages("GGally", repos = "http://cran.us.r-project.org")
if(!require(plotly)) install.packages("plotly", repos = "http://cran.us.r-project.org")
if(!require(maps)) install.packages("maps", repos = "http://cran.us.r-project.org")
if(!require(mapproj)) install.packages("mapproj", repos = "http://cran.us.r-project.org")
if(!require(car)) install.packages("car", repos = "http://cran.us.r-project.org")
if(!require(ggrepel)) install.packages("ggrepel", repos = "http://cran.us.r-project.org")
if(!require(stargazer)) install.packages("stargazer", repos = "http://cran.us.r-project.org")

# Load in the Future50 dataset & definitions ---------------------

Future50_dat <- read_csv(file = 'https://raw.githubusercontent.com/Jyednock/Future50/main/CaseStudy_Future50_data.csv')
Future50_def <- read_csv(file = 'https://raw.githubusercontent.com/Jyednock/Future50/main/CaseStudy_Future50_definitions.csv')


```

# Executuve Summary

This document presents an analysis of the Future50 dataset downloaded from www.restaurantbusinessonline.com on January 30, 2021. - The data set contains 50 rows and 9 columns. 

The purpose of this document is to provide some key highlights within the data, along with a one-pager report.

## Initial Data Exploration & Insights

The initial data exploration began with analyzing the data set and metric definitions. The first 5 rows of the dataset in addition to the metric definitions are listed below for reference:

```{r include=False}
Future50_dat %>%
  head(3)
```

```{r SummaryTable}
Future50_dat %>%
  select(Sales,YOY_Sales,Units,YOY_Units,Unit_Volume) %>%
  as.data.frame() %>%
  stargazer(type = "text")
```



```{r CorrelationMatrix,include=False,fig.height=6}
Future50_dat %>%
  select(Sales,YOY_Sales,Units,YOY_Units,Unit_Volume,Franchising) %>%
  ggpairs(aes(color = Franchising,fill = Franchising,alpha = .4))+
  theme(axis.text = element_text(size = 7))
```




#### Initial Insights

Based upon the dataset, it is apparent the restaurant ranking is based upon the YOY_Sales metric, as defined above. - This can be seen below:

```{r RankVsSalesIncrease,fig.height=5, message=FALSE,warning=FALSE}

# Incorporate gradient from white to green

Future50_dat %>%
  ggplot(aes(Rank,YOY_Sales))+
  geom_segment(aes(x = Rank,xend = Rank,y = 0,yend = YOY_Sales))+
  geom_point()+
 # geom_bar(stat = "identity",alpha = .4,color = "white",aes(fill = YOY_Sales),show.legend = F)+
  scale_y_continuous("YOY Sales Increase %",labels = percent)+
  geom_text(aes(label = ifelse(Rank %in% c(1,25,50),paste(round(100*YOY_Sales,0),"%",sep = ""),NA)),vjust = -.5,size = 3,fontface = 2)+
  labs(title = "YOY Sales Increase % Vs Restaurant Rank",
       subtitle = "*Rank determined by YOY_Sales; Ranks 1,25 and 50 are labeled")+
  theme_minimal()+
  scale_fill_gradient(low = "#B7FFBF",high = "#00AB08")
  #theme(panel.grid = element_blank())
```
We can see the highest ranked restaurants had some pretty tremendous year on year sales growth compared to the rest. - Why is this? Is there something in the dataset that can explain why restaurants achieved their year on year growth?


### YOY Sales Analysis

One of the metrics I wanted to dive into was the Sales percentage increase for 2019. - It might be interesting to find a relationship among the other metrics than can help explain the increase in sales within these restaurants.

```{r YOY Sales Increase,fig.height=4, message=FALSE,warning=FALSE}



lm_fit <- lm(YOY_Sales ~ YOY_Units,data = Future50_dat)
lm_outliers <- qqPlot(lm_fit,labels=row.names(Future50_dat), id.method="identify",
         simulate=TRUE, main="Q-Q Plot")


Future50_dat %>%
  ggplot(aes(YOY_Units,YOY_Sales))+
  geom_point(aes(color = Franchising),size = 4,alpha = .6)+
  scale_color_manual(values = c("Yes" = "gold","No" = "dodgerblue"))+
  theme_minimal()+
  geom_smooth(method = "lm",formula = y~x,se = F,color = "red",linetype = 2)+
  geom_text(aes(label = ifelse(Rank %in% lm_outliers,Restaurant,NA)),vjust = -.75)+
  scale_y_continuous("YOY Sales Increase %",labels = percent)+
  scale_x_continuous("YOY Premise Increase %",labels = percent)+
  labs(title = "YOY Sales Increase % Vs YOY Premises Incease %",
       subtitle = "*Red line represents linear regression line")




```

In general, there is a strong linear relationship between YOY_Units and YOY_Sales. - However, there are two outliers here that may be worth digging into. Playa bowls is over performing and Teriyaki Madness is underperforming considering the YOY_Units.



### Franchising Patterns

```{r FranchiseAnalysis, fig.height=4,fig.width=6}
Future50_dat %>%
  ggplot(aes(Franchising,Sales))+
  geom_boxplot()+
  geom_jitter()

Franchise_Count <- Future50_dat %>%
  group_by(Franchising) %>%
  summarise(Count = n())

# Franchise Vs Premises---

Future50_dat %>%
  left_join(Franchise_Count,by = "Franchising") %>%
  mutate(Count = paste0(Franchising, ' (N = ',Count, ') ')) %>%
  ggplot(aes(Count,Units))+
  geom_boxplot(alpha = .8)+
  geom_jitter(aes(color = Franchising),alpha = .7,size = 2)+
  labs(title = "Premises Vs Non-Franchised & Franchised Restaurants",
       x = "Franchising",y = "Premises")+
  scale_color_manual(values = c("Yes" = "gold","No" = "dodgerblue"))+
  theme_minimal()
  #geom_label(aes(data = Franchise_Count,label = Count))

# Franchise Vs Sales---

Future50_dat %>%
  left_join(Franchise_Count,by = "Franchising") %>%
  mutate(Count = paste0(Franchising, ' (N = ',Count, ') ')) %>%
  ggplot(aes(Count,Sales))+
  geom_boxplot(alpha = .8)+
  geom_jitter(aes(color = Franchising),alpha = .7,size = 2)+
  labs(title = "Sales Vs Non-Franchised & Franchised Restaurants",
       x = "Franchising",y = "Sales (Millions)")+
  scale_color_manual(values = c("Yes" = "gold","No" = "dodgerblue"))+
  scale_y_continuous(labels = dollar)+
  theme_minimal()

# Franchise Vs YOY Unit_Volume---

Future50_dat %>%
  left_join(Franchise_Count,by = "Franchising") %>%
  mutate(Count = paste0(Franchising, ' (N = ',Count, ') ')) %>%
  ggplot(aes(Count,Unit_Volume))+
  geom_boxplot(alpha = .8)+
  geom_jitter(aes(color = Franchising),alpha = .7,size = 2)+
  labs(title = "Average Premise Volume ($000) Vs Non-Franchised & Franchised Restaurants",
       x = "Franchising",y = "Average Premise Volume ($000)")+
  scale_color_manual(values = c("Yes" = "gold","No" = "dodgerblue"))+
  scale_y_continuous(labels = dollar)+
  theme_minimal()

  


```






```{r AveragePremiseVolume}


Future50_dat %>%
  ggplot(aes(Units,Unit_Volume))+
 # geom_point(size = 4,alpha = .5)+
  geom_point(aes(color = Franchising),size = 4,alpha = .5)+
  scale_color_manual(values = c("Yes" = "gold","No" = "dodgerblue"))+
  scale_y_continuous(labels = dollar)+
  labs(title = "Average Premise Volume ($000) Vs Premise Count",
       subtitle = "*Premise counts range from 7 to 105",
       x = "Premises",y = "Average Premise Volume ($000)")+
  theme_minimal()
```
There exists a non-linear relationship among premise count and average premise volume.


## Metric Engineering

### City & State

After a first glance of the dataset, I can see some new features that may be worth adding here. - One of them being the city and State location of these restaurants. It looks like this information can be extracted from the location metric with a comma delimiter. This may be helpful for geographically analyzing these top retaurants.

After creating city and state metrics, we can see how the top 50 retaurants are distributed across the states:

```{r StateMap,include=FALSE,warning=FALSE,message=FALSE}

Future50_dat <- Future50_dat %>%
  separate(Location,into = c("City","State"),remove = F,sep = ",")


Future50_dat <- Future50_dat %>%
  mutate(State_abbr = trimws(State),
         State_abbr = str_replace(State_abbr,'\\.',""),
         State_abbr = str_sub(State_abbr,start = 1,end = 2),
         State_abbr = toupper(State_abbr),
         State_abbr = ifelse(str_detect(State,"Texas"),"TX",
                             ifelse(str_detect(State,"Ariz"),"AZ",
                                    ifelse(str_detect(State,"D.C"),"MD",
                                           ifelse(str_detect(State,"Tenn"),"TN",State_abbr)))))

# TE to TX; AR to AZ; DC to MD; TE to TN

StateCount <- Future50_dat %>%
  left_join(statepop[,c("abbr","full")],by = c("State_abbr" = "abbr")) %>%
  group_by(State_abbr,full) %>%
  summarise(Count = n()) %>%
  ungroup() %>%
  mutate(full = tolower(full)) %>%
  rename(region = full)


states <- map_data("state")


snames <- data.frame(region=tolower(state.name), long_center=state.center$x, lat_center=state.center$y)
snames <- snames %>%
  left_join(StateCount,by = "region")
  

StateCount %>%
 # left_join(snames,by = "region") %>%
  merge(states,by = "region",all.y = T) %>%
  arrange(order) %>% 
  ggplot(aes(long,lat,group = group))+
  geom_polygon(aes(fill = Count))+
  geom_path()+
  theme_minimal()+
  geom_text(data=snames, aes(long_center, lat_center, group = NA,label=Count))+
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+
  coord_map()+
  labs(title = "State Distribution of Top 50 Ranked Retaurants")+
  theme(panel.grid = element_blank(),axis.title = element_blank(),axis.text = element_blank())
  


```

What I find interesting in the dataset is that the majority of the most populated states contain at least one of the top 50 retaurants, except for Illinois.

### Previous Year Sales and Units

Adding in the previous year's sales and units, we can do some one-to-one correlation analysis

```{r ExtraVars,include=F}


Future50_dat <- Future50_dat %>%
  mutate(Sales_prevyear = round(Sales/(1+YOY_Sales),0),
         Units_prevyear = round(Units/(1+YOY_Units),0),
         Sales_Grwth = Sales - Sales_prevyear,
         Units_Grwth = Units - Units_prevyear)

Future50_dat %>%
  ggplot(aes(Sales_prevyear,Sales,color = YOY_Units))+
  geom_point()
  

Saleslm <- lm(Sales ~ Sales_prevyear + YOY_Units,data = Future50_dat)
summary(Saleslm)



```






